install package
```{r}
options(download.file.method="libcurl")
install.packages("GGally")
```
```{r}
options(download.file.method="libcurl")
install.packages("RStoolbox")
```
```{r}
install.packages("plotly")
install.packages("htmlwidgets")
```
```{r}
install.packages("rstatix")
```

```{r}
library("RStoolbox")
```
library package
```{r}
library(sp)
library(raster)
library(rgeos)
library(rgdal)
library(rasterVis)
library(ggplot2)
library(terra)
library(sf)
library(stringr)
library(raster)
library(fs)
library(sf)
library(tidyverse)
```
List raster files excluding band 8 using the patter argument
Load raster layers into a stack
```{r}
listlandsat<-dir_info(here::here("Landsat")) %>%
  dplyr::filter(str_detect(path, "[B123456790].TIF")) %>%
  dplyr::select(path) %>%
  pull() %>%
  as.character() %>%
  stack()
```
Load the manchester boundary
```{r}
manchester_boundary <- st_read(here::here("Manchester_boundary",
                                          "Manchester_boundary.shp"))
crs(manchester_boundary)
crs(listlandsat)
```
There is an error with this dataset as band 8 does not fully align with the extent of the other raster layers.
get band 8
```{r}
b8list<-dir_info(here::here("Landsat"))%>%
  dplyr::filter(str_detect(path, "[B8].tif")) %>%
  dplyr::select(path)%>%
  pull()%>%
  as.character()%>%
  raster()
```
ngb is a nearest neighbour sampling method
and write out the resampled raster layer
```{r}
b8correct <- b8list%>%
  resample(., listlandsat$LC08_L1TP_203023_20190513_20190521_01_T1_B1, 
           method = "ngb") %>%
  writeRaster(.,str_c(here::here("Landsat"), 
                      names(b8list), 
                      sep="/"),
              format='GTiff', 
              overwrite=TRUE)
```
Load band 8 and add it to our raster stack
```{r}
b8backin<-dir_info(here::here("Landsat"))%>%
  dplyr::filter(str_detect(path, "[B8].tif")) %>%
  dplyr::select(path)%>%
  pull()%>%
  as.character()%>%
  raster()
listlandsat <- listlandsat %>%
  addLayer(., b8backin)
```
compare to see if both rasters have the same extent, number of rows and columns, projection, resolution and origin
```{r}
raster::compareRaster(listlandsat$LC08_L1TP_203023_20190513_20190521_01_T1_B1,
                      listlandsat$LC08_L1TP_203023_20190513_20190521_01_T1_B8)
```
create mask raster which cover only manchester boundry
now crop our temp data to the extent
```{r}
lsatmask <- listlandsat %>%
  raster::crop(.,manchester_boundary) %>%
  raster::mask(.,  manchester_boundary)
```
change filename in raster stack and write out .tiff file
```{r}
names(lsatmask) <- names(lsatmask)%>%
  str_c(., 
        "mask", 
        sep="_")
outputfilenames <-
  str_c("Landsat/", "mask/", names(lsatmask) ,sep="")
lsatmask %>%
  writeRaster(., outputfilenames, 
              bylayer=TRUE, 
              format='GTiff', 
              overwrite=TRUE)
```
load data by read in file and name the band
如果需要的raster少就用手动输入
manc<-stack(lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B1_mask）
```{r}
manc <- dir_info(here::here("Landsat","mask")) %>%
  dplyr::filter(str_detect(path, "[B1234567]_mask_mask_mask.tif")) %>%
  dplyr::filter(str_detect(path, "B11", negate=TRUE)) %>%
  dplyr::select(path) %>%
  pull() %>%
  stack()
names(manc) <- c('ultra-blue', 'blue', 'green', 'red', 'NIR', 'SWIR1', 'SWIR2')
```
extract specific information from a raster stack can use
```{r}
crs(manc) # projection
extent(manc) # extent
ncell(manc) # number of cells
dim(manc) # number of rows, columns, layers
nlayers(manc) # number of layers
res(manc) # xres, yres
```
plot data
in true colour (how humans see the world) and then false colour composites (using any other bands but not the combination of red, green and blue)
```{r}
manc_rgb <- stack(manc$red, manc$green, manc$blue)
manc_false <- stack(manc$NIR, manc$red, manc$green)
manc_rgb %>%
  plotRGB(.,axes=TRUE, stretch="lin")
manc_false %>%
    plotRGB(.,axes=TRUE, stretch="lin")
```
look at signle bands
```{r}
plot(manc$SWIR2)
```
check the similarity between bands
set the plot window size (2 by 2)
```{r}
par(mfrow = c(2,2))
plot(manc$blue, main = "Blue")
plot(manc$green, main = "Green")
plot(manc$red, main = "Red")
plot(manc$NIR, main = "NIR")
```
Look at the stats of these bands
```{r}
pairs(manc[[1:7]])
```
ggplot and GGally can make this look better
```{r}
library(ggplot2)
library(GGally)
manc %>%
  terra::as.data.frame(., na.rm=TRUE) %>%
  dplyr::sample_n(., 100) %>%
  ggpairs(.,axisLabels="none")
```
Raster calculation
compute temperature from raster data
first need to generate additional raster layer NDVI
Live green vegetation can be represented with the NIR and Red Bands through the normalised difference vegetation index (NDVI) as chlorophyll reflects in the NIR wavelength, but absorbs in the Red wavelength.
NDVI=(NIR-Red)/(NIR+Red)
can use function to repeat and process different data
```{r}
myfunction <- function(arg1, arg2, ... ){
  statements
  return(object)
}
```
use NDVI
```{r}
NDVIfun <- function(NIR, Red){
  NDVI <- (NIR - Red) / (NIR + Red)
  return(NDVI)
}
```
To be able to use this function throughout our analysis either copy it into the console or make a new R script, save it in your project then call it within this code using the source() function
```{r}
# source('prac8')
```
use NDVI function
```{r}
ndvi <- NDVIfun(manc$NIR, manc$red)
```
check output
```{r}
ndvi %>%
  plot(.,col = rev(terrain.colors(10)), main = "Landsat-NDVI")
```
look histogram
```{r}
ndvi %>%
  hist(., breaks = 40, main = "NDVI Histogram", xlim = c(-.3,.8))
```
can reclassify to the raster to show use what is most likely going to vegetation based on the histogram using the 3rd quartile — anything above the 3rd quartile we assume is vegetation
this is an assumption for demonstration purposes, if you were to do something similar in your assignment be sure to provide reasoning with linkage to literature (e.g. policy or academic)
0.3 is assumption
```{r}
veg <- ndvi %>%
  reclassify(., cbind(-Inf, 0.3, NA))
veg %>%
  plot(.,main = 'Possible Veg cover')
```
look at this in relation to Manchester as a whole
```{r}
manc_rgb %>%
  plotRGB(.,axes = TRUE, stretch = "lin", main = "Landsat True Color Composite")
veg %>%
  plot(., add=TRUE, legend=FALSE)
```
The goal of this final section is to set up a mini investigation to see if there is a relationship between urban area and temperature. If our hypothesis is that there is a relationship then our null is that there is not a relationship
# Calcualte the Top of Atmopshere (TOA) spectral radiance from the Digital Number (DN) using
# λ = Grescale * QCAL + Brescale
TOA spectral radiance is light reflected off the Earth as seen from the satellite measure in radiance units.
Grescale and Brescale represent the gain and bias of the image
with QCAL the Digital Number (DN) — how the raw Landsat image is captured. 
To go from DN to spectral radiance we use the calibration curve, created before the launch of the sensor. 
Bias is the spectral radiance of the sensor for a DN of 0, Gain is the gradient of the slope for other values of DN.
Grescale and Brescale are available from the .MTL file provided when you downloaded the Landsat data. Either open this file in notepad and extract the required values for band 10 gain (MULT_BAND) and bias (ADD_BAND) Or we can automate it using the MTL() function within the RStoolbox package
```{r}
MTL<-dir_info(here::here("Landsat")) %>%
  dplyr::filter(str_detect(path, "MTL.txt")) %>%
  dplyr::select(path)%>%
  pull()%>%
  readMeta()
head(MTL)
```
extract value from MTL for band 10
```{r}
offsetandgain <-MTL %>%
  getMeta("B10_dn", metaData = ., what = "CALRAD")
offsetandgain
```
Run the calculation using the band 10 raster layer
```{r}
TOA <- offsetandgain$gain *
  lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B10_mask + 
  offsetandgain$offset
```
convert the TOA to Brightness Temperature Tb using the following equation
Tb = K2 / ln((K1/λ)+1)
Brightness temperature is the radiance travelling upward from the top of the atmosphere to the satellite in units of the temperature of an equivalent black body.
K1 (774.8853) and K2 (1321.0789) are pre launch calibration constants provided by USGS.
these can extract from MTL
```{r}
Calidata <- MTL$CALBT %>%
  terra::as.data.frame() %>%
  mutate(Band=rownames(.)) %>%
  filter(Band=="B10_dn")
K1 <- Calidata %>%
  dplyr::select(K1) %>%
  pull()
K2 <- Calidata %>%
  dplyr::select(K2) %>%
  pull()
Brighttemp <- (K2 / log((K1 / TOA) + 1))
```
determine emissivity of each pixel by using NDVI and Tb
first calculate fractional vegetation of each pixel
Fv = ((NDVI-NDVImin)/(NDVImax-NDVImin))^2
Here, NDVImin is the minimum NDVI value (0.2) where pixels are considered bare earth and  
NDVImax is the value at which pixels are considered healthy vegetation (0.5).
```{r}
facveg <- (ndvi-0.2/0.5-0.2)^2
```
compute emissivity
ε = 0.004 * Fv + 0.986
Emissivity is the ratio absorbed radiation energy to total incoming radiation engery compared to a blackbody (which would absorb everything), being ab measure of absoptivity.
```{r}
emiss <- 0.004*facveg+0.986
```
compute land surface temperature
LST = Tb / (1+(λϱTb/(p))lnε)
p = h*c/ϱ
h which is Plank’s constant 6.626×10^(−34)Js
c which is the velocity of light in a vaccum 2.998 × 10^8 m/sec
ϱ which is the Boltzmann constant of 1.38 × 10^(-23) J/K
```{r}
Boltzmann <- 1.38*10e-23
Plank <- 6.626*10e-34
c <- 2.998*10e8
p <- Plank*(c/Boltzmann)
```
λ which is the effective wavelength of our data (10.9 for Landsat 8 band 10)
ε emissivity
Tb brightness Temperature
```{r}
lambda <- 1.09e-5
LST <- Brighttemp/(1 +(lambda*Brighttemp/p)*log(emiss))
LST
```
High value because in Kevlin not degrees Celcius
```{r}
LST <- LST-273.15
plot(LST)
```
Calculate urban area from Landsat data
use the Normalized Difference Built-up Index (NDBI) algorithm for identification of built up regions using the reflective bands: Red, Near-Infrared (NIR) and Mid-Infrared (MIR)
NDBI = (ShortwaveInfrared(SWIR)-NearInfrared(NIR))/(ShortwaveInfrared(SWIR)+NearInfrared(NIR))
SWIR is band 6 and the NIR band 5
```{r}
NDBI=((lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B6_mask-
         lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B5_mask)/
        (lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B6_mask+
        lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B5_mask))
plot(NDBI)
```
or NDVIfun
```{r}
# NDBIfun <- NDVIfun(lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B6_mask,lsatmask$LC08_L1TP_203023_20190513_20190521_01_T1_B5_mask)
```
urban area and temperature relationship
```{r}
plot(values(NDBI), values(LST))
```
a lot of points overlap problem so take random subset of same pixels from both raster layers
stack layers
```{r}
computeddata <- LST %>%
  stack(.,NDBI) %>%
  terra::as.data.frame() %>%
  na.omit() %>%
  dplyr::sample_n(., 500) %>%
  dplyr::rename(Temp="layer.1", NDBI="layer.2")
plot(computeddata$Temp, computateddata$NDBI)
```
jazz things up load some more packages
```{r}
library(plotly)
library(htmlwidgets)
```
Transfrom the data to a data.frame to work with ggplot, then plot
```{r}
heat<-ggplot(computeddata, aes(x = Temp, y = NDBI))+
  geom_point(alpha=2, colour = "#51A0D5")+
  labs(x = "Temperature", 
       y = "Urban index",
       title = "Manchester urban and temperature relationship")+
   geom_smooth(method='lm', se=FALSE)+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))
# interactive plot
ggplotly(heat)
```
how about whole dataset
use hexbin
```{r}
computeddatafull <- LST %>%
  stack(.,NDBI)%>%
  terra::as.data.frame() %>%
  na.omit() %>%
  dplyr::rename(Temp="layer.1", NDBI="layer.2")

hexbins <- ggplot(computeddatafull, 
                  aes(x=Temp, y=NDBI)) +
  geom_hex(bins=100, na.rm=TRUE) +
  labs(fill = "Count per bin")+
  geom_smooth(method='lm', se=FALSE, size=0.6)+
  theme_bw()

ggplotly(hexbins)
```
statistical summary
```{r}
library(rstatix)
Correlation <- computeddatafull %>%
  cor_test(Temp, NDBI, use = "complete.obs", method = c("pearson"))
Correlation
```
p-value: tells us whether there is a statistically significant correlation between the datasets and if that we can reject the null hypothesis if p<0.05 (there is a 95% chance that the relationship is real).
cor: Product moment correlation coefficient
conf.low and con.high intervals: 95% confident that the population correlation coeffieicent is within this interval
statistic value (or t, or test statistic)

work critical t value
0.05 is the confidence level (95%)
2 means a 2 sided test
198268 is the degrees of freedom (df), being the number of values we have -2
```{r}
computeddatafull %>%
  pull(Temp) %>%
  length()
```
```{r}
length(computeddatafull)
```
```{r}
abs(qt(0.05/2, 198268))
```
Here, as our t values is > than the critical value we can say that there is a relationship between the datasets. However, we would normally report the p-value…which we can get from..

As p<0.05 is shows that are variables are have a statistically significant correlation… so as urban area (assuming the index in representative) per pixel increases so does temperature…therefore we can reject our null hypothesis… but remember that this does not imply causation!!

LSOA/MSOA
```{r}
UK_LSOA <- st_read(here::here("LSOA",
                           "Lower_Layer_Super_Output_Areas_December_2011_Generalised_Clipped__Boundaries_in_England_and_Wales.shp")) %>%
  st_transform(., 32630)

```
```{r}
MSOA <- st_read(here::here("MSOA_EngWal_Dec_2011_Generalised_ClippedEW_0",
                           "Middle_Layer_Super_Output_Areas_December_2011_Generalised_Clipped_Boundaries_in_England_and_Wales.shp")) %>%
   st_transform(., 32630)
```
select only MSOA within boundary
```{r}
manchester_MSOA <- MSOA[manchester_boundary, , op=st_within]
```
select only LSOA that intersect MSOA
```{r}
manchester_LSOA <- UK_LSOA[manchester_MSOA,]
```
extraction with raster::extract(). fun() specifies how to summarise the pixels within the spatial unit (LSOA), na.rm()=TRUE ignores NA values and df=TRUE outputs the result to a dataframe.
```{r}
# extract mean LST value per LSOA
LST_per_LSOA <- raster::extract(LST, manchester_LSOA, fun=mean, na.rm=TRUE, df=TRUE)
# add the LSOA ID back
LST_per_LSOA$objectid <- manchester_LSOA$objectid
# join the average temp to the sf
manchester_LSOA_temp <- manchester_LSOA %>%
  left_join(.,
            LST_per_LSOA,
            by="objectid")%>%
  dplyr::rename(temp=layer)
```
define urban as NDBI greater than 0
Sum the pixels that are grater than 0 per LSOA
count the pixels per LSOA
```{r}
NDBI_urban<- NDBI > 0
NDBI_urban_per_LSOA <- raster::extract(NDBI_urban, manchester_LSOA, fun=sum, na.rm=TRUE, df=TRUE)
NDBI_per_LSOA_cells <- raster::extract(NDBI_urban, manchester_LSOA, na.rm=TRUE, df=TRUE, cellnumbers=TRUE)
NDBI_per_LSOA2_cells<- NDBI_per_LSOA_cells %>%
  count(ID)
```
add the LSOA ID to the urban area
add the LSOA ID to the number of cells
```{r}
NDBI_urban_per_LSOA$objectid <- manchester_LSOA$objectid
NDBI_per_LSOA2_cells$objectid <- manchester_LSOA$objectid
```
join these two
```{r}
urban_info_LSOA <- NDBI_urban_per_LSOA %>%
  left_join(.,
            NDBI_per_LSOA2_cells,
            by = "ID")
urban_info_LSOA
```
remove what you don't need and rename
```{r}
Urban_info_LSOA_core_needed <- urban_info_LSOA %>%
  dplyr::select(.,
                layer,
                n,
                ID) %>%
  dplyr::mutate(percent_urban=layer/n*100)
Urban_info_LSOA_core_needed
```
join the data 
one sf with temp and % urban per LSOA
```{r}
manchester_LSOA_temp_urban <- manchester_LSOA_temp %>%
  left_join(.,
             Urban_info_LSOA_core_needed,
             by="ID")
```
map both temperature (and the % of urban area) within a LSOA individually
In our map we want to include some place names from Open Street Map
```{r}
Places <- st_read(here::here("greater-manchester-latest-free.shp",
                                          "gis_osm_places_free_1.shp")) %>%
   st_transform(., 32630)
plot(Places)
```

```{r}
manchester_Places <- Places[manchester_boundary,] %>%
  filter(fclass=="city")
```
this first bit makes the box bigger
so we can have a north arrow not overlapping the data
```{r}
bbox_new <- st_bbox(manchester_LSOA_temp_urban)
yrange <- bbox_new$ymax - bbox_new$ymin
bbox_new[4] <- bbox_new[4] + (0.1 * yrange)
bbox_new[2] <- bbox_new[2] - (0.1 * yrange)
```

```{r}
library(tmap)
tmap_mode("plot")
tm1 <- tm_shape(manchester_LSOA_temp_urban, bbox = bbox_new) + 
  tm_polygons("temp",
              palette="OrRd",
              legend.hist=TRUE,
              title="Temperature")+
  tm_shape(manchester_Places, bbox=bbox_new)+
  tm_dots(size=0.1, col="white")+
  tm_text(text="name", size=0.75, ymod=-0.5, col="white", fontface = "bold")+
  #tm_legend(show=FALSE)+
  tm_layout(frame=FALSE,
            legend.outside=TRUE)+
  tm_compass(type = "arrow", size=1, position = c("left", "top")) +
  tm_scale_bar(position= c("left", "bottom"), breaks=c(0,2,4), text.size = .75)
tm1
```

